---
title: "ADA Final Project Analysis"
author: "Taylor Kaser"
date: "2022-11-20"
output: html_document
---

```{r include=FALSE}
#open libraries
pacman::p_load(haven, tidyverse, naniar, VIM, mice, lattice, table1, lubridate, lmtest) 
```


```{r include=FALSE}
# Read in the data
setwd("C:/Users/15416/OneDrive/Documents/Fall 2022/ADA/Final Project/ICPSR_38429")

# Reading in the data from an external file as opposed to using the dataframe created in the data management Rmd
# reduces memory usage needed for imputations
victim_data <- read_csv("Final_project_data.csv")

# Change data types from characters to factors
victim_data <- victim_data %>%
  mutate_if(is.character, as.factor) %>%
  # Recorder factor levels
  mutate(bio_sex = factor(bio_sex,
                       levels = c("Male","Female"))) %>%
  mutate(race = factor(race,
                       levels = c("White", "Black",
                                      "Other"))) %>%
  mutate(knows_perp = factor(knows_perp,
                       levels = c('Stranger','Relative','Other Known','Intimate Partner'))) %>%
  mutate(repeatinj_yn = factor(repeatinj_yn,
                       levels = c('No','Yes'))) %>%
  mutate(age.times.logage = age * log(age)) # create term to test linearity assumption for logistic   regression
```


# Statistical analysis 

## Missing data

Plan: Missing data for any variables included in the analysis will be addressed using multiple imputation. 

```{r Examine missing data patterns}
# Examine missing data patterns
md.pattern(victim_data, rotate.names=TRUE)

# 250/337 appear to be complete (74.2%)
## >10% missing. Justifies need to perform multiple imputations
# There does not appear to be strong correlations between variables in terms of missingness
# Total of 6 missing variable patterns - only 2 include more than one missing variable

# Determine percentage of missing data in each column.
colMeans(is.na(victim_data))*100
# data for relationship to perpetrator is >10%. Also want to do MI because we have
# a small sample size and want to retain as many observations as possible
```

```{r MI}
# Impute missing data
# Setting m=9 since there are 9 variables that will be included in the imputation algorithm

imp<-mice(victim_data, m=9, maxit=5, seed=219) 

# Remove variables we do not want to use as predictors in the imputation algorithm:
# Participant and household ID, year of survey, and month-year of incident (only used for identifying the index injury, not as a predictor of the outcome)
pred <- imp$predictorMatrix
pred[, "IDPER"] <- 0
pred[, "YEARQ"] <- 0
pred[, "IDHH"] <- 0
pred[, "my_incident"] <- 0
pred[, "age.times.logage"] <- 0
pred

imp<-mice(victim_data, pred = pred, m=9, maxit=5, seed=219)

# print initial imputation details
imp

```

```{r MI diagnostics}
# Run diagnostics for the imputed model
## the rows contain the imputed values and the columns are the multiple imputations

# check relationship to perpetrator
imp$imp$knows_perp

# check law enforcement involvement
imp$imp$report_LE

# check injury type
imp$imp$inj_type

# check incident location (near or at participants' residence (Y/N))
imp$imp$incdt_location

# # Column 1 are observation numbers
# # The remaining columns refer to the imputation number
```

```{r Obtain complete imputed dataset}

# Obtain complete imputed datasets in a dataframe using the complete function from the mice package
## We will use this for testing regression assumptions

imp_data1  <-mice::complete(data=imp, action = 1) 
imp_data2  <-mice::complete(data=imp, action = 2) 
imp_data3  <-mice::complete(data=imp, action = 3) 
imp_data4  <-mice::complete(data=imp, action = 4) 
imp_data5  <-mice::complete(data=imp, action = 5)  
imp_data6  <-mice::complete(data=imp, action = 6)  
imp_data7  <-mice::complete(data=imp, action = 7) 
imp_data8  <-mice::complete(data=imp, action = 8)  
imp_data9  <-mice::complete(data=imp, action = 9)  

#longimp <- mice::complete(data=imp, action="long") # To get all the imputations in one dataset provide the action "long". Two columns are added: 1) .imp, integer, referring the imputation number, and 2) .id, character, the row names of data$data;

# Compare imputed dataset 1 with observational data

table1(~ age + bio_sex + race + knows_perp + report_LE + inj_type + incdt_location + sought_care | repeatinj_yn,
       data = victim_data)

table1(~ age + bio_sex + race + knows_perp + report_LE + inj_type + incdt_location + sought_care | repeatinj_yn,
       data = imp_data1)

```

## Create bar plot for the association between seeking medical care and violent reinjury

```{r}
# Use a grouped bar chart
victim_data %>%
  group_by(sought_care, repeatinj_yn) %>%
  count() %>%
  group_by(sought_care) %>%
  mutate(perc.sought_care = 100*n/sum(n)) %>%
  ggplot(aes(x = sought_care, y = perc.sought_care, fill = repeatinj_yn)) +
  geom_col(position = "stack") +
  theme_minimal(base_size = 12) +
  labs(x = "\nSought care after violent injury?",
       y = "Percent of participants\n",
       title = "Figure 1: The association between seeking care for violent injury \nand experiencing repeat violent injury (Observational Data).",
       fill = "Experienced \nviolent reinjury?") +
  scale_fill_manual(values = c("#2BC9B5", "#A0345C"))  

imp_data1 %>%
  group_by(sought_care, repeatinj_yn) %>%
  count() %>%
  group_by(sought_care) %>%
  mutate(perc.sought_care = 100*n/sum(n)) %>%
  ggplot(aes(x = sought_care, y = perc.sought_care, fill = repeatinj_yn)) +
  geom_col(position = "stack") +
  theme_minimal(base_size = 12) +
  labs(x = "\nSought care after violent injury?",
       y = "Percent of participants\n",
       title = "Figure 2: The association between seeking care for violent injury \nand experiencing repeat violent injury (Imputed Dataset #1).",
       fill = "Experienced \nviolent reinjury?") +
  scale_fill_manual(values = c("#2BC9B5", "#A0345C"))  

imp_data2 %>%
  group_by(sought_care, repeatinj_yn) %>%
  count() %>%
  group_by(sought_care) %>%
  mutate(perc.sought_care = 100*n/sum(n)) %>%
  ggplot(aes(x = sought_care, y = perc.sought_care, fill = repeatinj_yn)) +
  geom_col(position = "stack") +
  theme_minimal(base_size = 12) +
  labs(x = "\nSought care after violent injury?",
       y = "Percent of participants\n",
       title = "Figure 3: The association between seeking care for violent injury \nand experiencing repeat violent injury (Imputed Dataset #2).",
       fill = "Experienced \nviolent reinjury?") +
  scale_fill_manual(values = c("#2BC9B5", "#A0345C")) 

# Might not include in the final presentation. These aren't that helpful.

```

##  Primary objective: Determine whether accessing healthcare following a violent injury is associated with lower odds of violent reinjury.

```{r fit logistic regression model}
# Fit logistic regression model for violent reinjury

reinj_model <- glm(repeatinj_yn ~ age + bio_sex + race + knows_perp + report_LE +
                   inj_type + incdt_location + sought_care,
                      data = victim_data,
                      family = binomial(link = "logit"))

# Fit logistic regression model for violent reinjury using imputed data

reinj_model_imp <- with(imp, glm(repeatinj_yn ~ age + bio_sex + race + knows_perp + report_LE +
                   inj_type + incdt_location + sought_care,
                      data = victim_data,
                      family = binomial(link = "logit")))

# examine model output
summary(reinj_model)
summary(pool(reinj_model_imp))

# Exponentiate/generate confidence intervals using tidy::broom
broom::tidy(reinj_model, exponentiate = TRUE, conf.int = TRUE)
broom::tidy(pool(reinj_model_imp), exponentiate = TRUE, conf.int = TRUE)

# Note: coefficient estimates are ~ the same for the imputed and non-imputed datasets
```

```{r Fit unadjusted models}

# Fit unadjusted models for comparison

reinj_model_unadj <- glm(repeatinj_yn ~ sought_care,
                      data = victim_data,
                      family = binomial(link = "logit"))

# Fit logistic regression model for violent reinjury using imputed data

reinj_model_impunadj <- with(imp, glm(repeatinj_yn ~ sought_care,
                      data = victim_data,
                      family = binomial(link = "logit")))

# examine model output
summary(reinj_model_unadj)
summary(pool(reinj_model_impunadj))

# Exponentiate/generate confidence intervals using tidy::broom
broom::tidy(reinj_model_unadj, exponentiate = TRUE, conf.int = TRUE)
broom::tidy(pool(reinj_model_impunadj), exponentiate = TRUE, conf.int = TRUE)

# Note: coefficient estimates are ~ the same for the imputed and non-imputed datasets
```


## Test assumptions

```{r Test Linearity}
# Test the linearity assumption for age using the Box Tidwell test. 

# We created a term for age*log(age) previously in order to prepare for testing this assumption.  
#Remember, a SIGNIFICANT (i.e. p < .05) coefficient means the assumption IS VIOLATED

# Test assumption on imputed dataset #1

reinj_model_bttest1 <- (glm(repeatinj_yn ~ age + bio_sex + race + knows_perp + report_LE +
                   inj_type + incdt_location + sought_care + age.times.logage,
                      data = imp_data1,
                      family = binomial(link = "logit")))

broom::tidy(reinj_model_bttest1, exponentiate = TRUE, conf.int = TRUE)
# p = 0.395 - assumption is met.

# Test assumption on imputed dataset #2

reinj_model_bttest2 <- (glm(repeatinj_yn ~ age + bio_sex + race + knows_perp + report_LE +
                   inj_type + incdt_location + sought_care + age.times.logage,
                      data = imp_data2,
                      family = binomial(link = "logit")))

broom::tidy(reinj_model_bttest2, exponentiate = TRUE, conf.int = TRUE)
# p = 0.546 - assumption is met.

# Test assumption on imputed dataset #3

reinj_model_bttest3 <- (glm(repeatinj_yn ~ age + bio_sex + race + knows_perp + report_LE +
                   inj_type + incdt_location + sought_care + age.times.logage,
                      data = imp_data3,
                      family = binomial(link = "logit")))

broom::tidy(reinj_model_bttest3, exponentiate = TRUE, conf.int = TRUE)
# p = 0.485 - assumption is met.

# Test assumption on imputed dataset #4

reinj_model_bttest4 <- (glm(repeatinj_yn ~ age + bio_sex + race + knows_perp + report_LE +
                   inj_type + incdt_location + sought_care + age.times.logage,
                      data = imp_data4,
                      family = binomial(link = "logit")))

broom::tidy(reinj_model_bttest4, exponentiate = TRUE, conf.int = TRUE)
# p = 0.457 - assumption is met.

# Test assumption on imputed dataset #5

reinj_model_bttest5 <- (glm(repeatinj_yn ~ age + bio_sex + race + knows_perp + report_LE +
                   inj_type + incdt_location + sought_care + age.times.logage,
                      data = imp_data5,
                      family = binomial(link = "logit")))

broom::tidy(reinj_model_bttest5, exponentiate = TRUE, conf.int = TRUE)
# p = 0.494 - assumption is met.

# Test assumption on imputed dataset #6

reinj_model_bttest6 <- (glm(repeatinj_yn ~ age + bio_sex + race + knows_perp + report_LE +
                   inj_type + incdt_location + sought_care + age.times.logage,
                      data = imp_data6,
                      family = binomial(link = "logit")))

broom::tidy(reinj_model_bttest6, exponentiate = TRUE, conf.int = TRUE)
# p = 0.543 - assumption is met.

# Test assumption on imputed dataset #7

reinj_model_bttest7 <- (glm(repeatinj_yn ~ age + bio_sex + race + knows_perp + report_LE +
                   inj_type + incdt_location + sought_care + age.times.logage,
                      data = imp_data7,
                      family = binomial(link = "logit")))

broom::tidy(reinj_model_bttest7, exponentiate = TRUE, conf.int = TRUE)
# p = 0.363 - assumption is met.

# Test assumption on imputed dataset #8

reinj_model_bttest8 <- (glm(repeatinj_yn ~ age + bio_sex + race + knows_perp + report_LE +
                   inj_type + incdt_location + sought_care + age.times.logage,
                      data = imp_data8,
                      family = binomial(link = "logit")))

broom::tidy(reinj_model_bttest8, exponentiate = TRUE, conf.int = TRUE)
# p = 0.683 - assumption is met.

# Test assumption on imputed dataset #9

reinj_model_bttest9 <- (glm(repeatinj_yn ~ age + bio_sex + race + knows_perp + report_LE +
                   inj_type + incdt_location + sought_care + age.times.logage,
                      data = imp_data9,
                      family = binomial(link = "logit")))

broom::tidy(reinj_model_bttest9, exponentiate = TRUE, conf.int = TRUE)
# p = 0.492 - assumption is met.

```

```{r Test Multicollinearity}
# Test multicollinearity assumption using Variance Inflation Factors

# We will create and test individual models for each imputed dataset

### IMPUTED MODEL 1

reinj_model_vif1 <- (glm(repeatinj_yn ~ age + bio_sex + race + knows_perp + 
                           report_LE + inj_type + incdt_location + sought_care,
                      data = imp_data1,
                      family = binomial(link = "logit")))

# Use the vif function from the car package

jtools::summ(reinj_model_vif1, vifs = TRUE)
# All VIF < 2.5. Assumption is met.

### IMPUTED MODEL 2

reinj_model_vif2 <- (glm(repeatinj_yn ~ age + bio_sex + race + knows_perp + 
                           report_LE + inj_type + incdt_location + sought_care,
                      data = imp_data2,
                      family = binomial(link = "logit")))

# Use the vif function from the car package

jtools::summ(reinj_model_vif2, vifs = TRUE)
# All VIF < 2.5. Assumption is met.

### IMPUTED MODEL 3

reinj_model_vif3 <- (glm(repeatinj_yn ~ age + bio_sex + race + knows_perp + 
                           report_LE + inj_type + incdt_location + sought_care,
                      data = imp_data3,
                      family = binomial(link = "logit")))

# Use the vif function from the car package

jtools::summ(reinj_model_vif3, vifs = TRUE)
# All VIF < 2.5. Assumption is met.

### IMPUTED MODEL 4

reinj_model_vif4 <- (glm(repeatinj_yn ~ age + bio_sex + race + knows_perp + 
                           report_LE + inj_type + incdt_location + sought_care,
                      data = imp_data4,
                      family = binomial(link = "logit")))

# Use the vif function from the car package

jtools::summ(reinj_model_vif4, vifs = TRUE)
# All VIF < 2.5. Assumption is met.

### IMPUTED MODEL 5

reinj_model_vif5 <- (glm(repeatinj_yn ~ age + bio_sex + race + knows_perp + 
                           report_LE + inj_type + incdt_location + sought_care,
                      data = imp_data5,
                      family = binomial(link = "logit")))

# Use the vif function from the car package

jtools::summ(reinj_model_vif5, vifs = TRUE)
# All VIF < 2.5. Assumption is met.

### IMPUTED MODEL 6

reinj_model_vif6 <- (glm(repeatinj_yn ~ age + bio_sex + race + knows_perp + 
                           report_LE + inj_type + incdt_location + sought_care,
                      data = imp_data6,
                      family = binomial(link = "logit")))

# Use the vif function from the car package

jtools::summ(reinj_model_vif6, vifs = TRUE)
# All VIF < 2.5. Assumption is met.

### IMPUTED MODEL 7

reinj_model_vif7 <- (glm(repeatinj_yn ~ age + bio_sex + race + knows_perp + 
                           report_LE + inj_type + incdt_location + sought_care,
                      data = imp_data7,
                      family = binomial(link = "logit")))

# Use the vif function from the car package

jtools::summ(reinj_model_vif7, vifs = TRUE)
# All VIF < 2.5. Assumption is met.

### IMPUTED MODEL 8

reinj_model_vif8 <- (glm(repeatinj_yn ~ age + bio_sex + race + knows_perp + 
                           report_LE + inj_type + incdt_location + sought_care,
                      data = imp_data8,
                      family = binomial(link = "logit")))

# Use the vif function from the car package

jtools::summ(reinj_model_vif8, vifs = TRUE)
# All VIF < 2.5. Assumption is met.

### IMPUTED MODEL 9

reinj_model_vif9 <- (glm(repeatinj_yn ~ age + bio_sex + race + knows_perp + 
                           report_LE + inj_type + incdt_location + sought_care,
                      data = imp_data9,
                      family = binomial(link = "logit")))

# Use the vif function from the car package

jtools::summ(reinj_model_vif9, vifs = TRUE)
# All VIF < 2.5. Assumption is met.
```

## Identify Influential Observations

```{r Plot Cooks Distance}
# Identify any extreme outliers
# Used model created with observational data

# Plot Cook's Distance - the higher the distance, the more it is considered an outlier
### Note, this works better for continuous variables than categorical
plot(reinj_model, which = 4, id.n = 3, col="red") 

```


```{r Cooks D calculations}
## Get Cook's D calculations for each observation and using the augment function  from Broom.

model.data <- broom::augment(reinj_model) %>%  # Allows you to get Cook's distance in a dataframe
  mutate(index = 1:n()) 

head(model.data)
```


```{r}
## Identify and exclude observations with a Cook's Distance greater than a cutoff, use 3 x the mean Cook's D

# Cutoffs from the lecture
#cuttoff <- mean(model.data$.cooksd)*3 - (removed 33.8% of data points)
# cuttoff <- (4/337) # defines cutoff variable (4/N) - (removed 35% of data points)
 cuttoff <- 1 # (removed 25.8% of the data)

# Define cuttoff using influential values revealed by the graph 
#cuttoff <- 0.04 - (removed 26.7% of data points) 

victim_data_out <- model.data %>% # excludes observations
  dplyr::filter(.cooksd < cuttoff)

nrow(victim_data_out)/nrow(victim_data) # check percent kept

# Elect to keep all observations since Cook's distance indicates removing 1/4 - 1/3 of the data.
```
## Secondary Objective: Determine whether injury type modifies the association between accessing healthcare following violent injury and violent reinjury

```{r}
# Compare models with and without interaction between injury type and accessing care
## Use D3 for LR test

LR_int_model <- with(imp,glm(repeatinj_yn ~ age + bio_sex + race + knows_perp + report_LE +
                    incdt_location + sought_care*inj_type,
                      data = victim_data,
                      family = binomial(link = "logit")))

D3(LR_int_model, reinj_model_imp)

# p-value = 0.451 - no effect modification by injury type

```







